Ludwig User Tutorial

### 2017-06-20 docs/tutorial/tutorial.tex
Edited tutorial section for test1
git-svn-id: https://ccpforge.cse.rl.ac.uk/svn/ludwig/trunk@3139 40876441-3496-0410-a613-a2fca35b2ae2

The system requirements are an up-to-date version of a C compiler as well as working installations 
of Apache's Subversion (SVN) version control system, the visualisation tool
ParaView (\hyperref[ParaView]{https://www.paraview.org}) and the MPI-library 
for compiling the parallel version of the code.

@@

>>>>

The system requirements are an up-to-date version of a C compiler as well as working installations 
of Apache's Subversion (SVN) version control system, the visualisation tool
ParaView (\hyperref[ParaView]{https://www.paraview.org}), the MPI-library 
for compiling the parallel version of the code and \texttt{gnuplot} for simple visualisations.

@@

\section{Tutorial Test Exercises}

\subsection{Test 1: Poiseuille Flow of a Simple Fluid in a Cavity}

In this section the Poiseuille flow of a simple fluid in a cavity, is outlined. 
The example is a step-by-step guide for performing a short run ($10,000$ time-steps) 
with \texttt{Ludwig} code, followed by a short introduction to the post-processing 
and data analysis procedures. Before starting, the code should be compiled in 
parallel and the executable \texttt{Ludwig.exe} should be copied into your working directory.

\begin{enumerate}
\item Copy the input file \texttt{input} located in the directory \texttt{/trunk/docs/tutorial/test1} 
to your working directory. Note that this input file is a modified version of the 
reference input file \texttt{input.ref} located in \texttt{/trunk/src} with a limited set of parameters
for this run. 
\item Run the code in your working directory by issuing the command:
\begin{lstlisting}
$ mpirun -np 4 ./Ludwig.exe input
\end{lstlisting}
\item You should now see statistics being reported on your standard output and 
output files being created with names of the form \texttt{vel-XXXXXXXX.001-001}. 
These files contain the components of the velocity vectors in Cartesian coordinates at each lattice point in the simulation. 
\item The first step to the post-processing of these data is to copy the files \texttt{vtk\_extract.c} 
and \texttt{vtk\_extract\_script.py} from \texttt{/trunk/util} into your working directory.
\item Edit \texttt{vtk\_extract.c}. Specifically, set the flags for:
\begin{enumerate}
\item \texttt{input\_binary\_=0} \\ if the velocity data is in ASCII format (see flag in input file),
\item \texttt{output\_binary\_=0} \\ as the output of the postprocessing routines should also be in ASCII format,
\item \texttt{output\_index=1} \\ to obtain Cartesian lattice indices \textit{i, j, k}, 
\item \texttt{vtk\_header=0} \\ to omit headers (these will be required later for visualisation in \textit{ParaView})
\item \texttt{output\_cmf=0} to produce data in row-major format (column-major format is required for \textit{ParaView}).
\end{enumerate}
\item Edit \texttt{vtk\_extract\_script.py}. Set:
\begin{enumerate}
\item \texttt{nstart=1000}, \texttt{nint=1000} and \texttt{nend=10000} to set the range and increments of the velocity data
to be processed.
\item \texttt{vel=1} for velocity post-processing, 
\item The other flags should be set to zero, e.g. \texttt{phi=0}. 
\end{enumerate}
\item You should now see files named \texttt{vel-XXXXXXXX.vtk} (Note: The extension \texttt{.vtk} is only used here for simplicity.
It does not mean the generated data is in vtk-format). 
The first three columns of these ASCII files are the spatial lattice indices followed by the Cartesian components of the velocity vectors.\\

To plot the velocity profile of the system, you can plot the $y$-component of the velocity against 
the spatial $x$-coordinate on the lattice with the command:
\begin{lstlisting}
gnuplot> p`vel-XXXXXXXX.vtk' u 1:5
\end{lstlisting} 
By plotting the data in each file, you should be able to see the velocity profile converging to the final profile 
as shown in Fig. \ref{fig:velocityProfile}. In the directory \texttt{/trunk/docs/tutorial/test1} a reference file of the final 
output at time step $10,000$ can be found, in addition to a file containing the standard output of the code.
\end{enumerate}

\begin{figure}[h]
\begin{center}
\includegraphics[width=0.6\linewidth]{velProf.png}
  \caption{The parabolic velocity profile of a simple fluid in a cavity after $10,000$ time-steps}.
  \label{fig:velocityProfile}
  \end{center}
\end{figure}
